Limited plasticity but increased variance in physiological rates across ectotherm populations under climate change

### 2024-07-13 ms.qmd
Added in detailed trait anlaysis for lnVR given we are focusing more on this for easier interpretation

---
title: "Physiological plasticity and variance in physiological rates vary across ectotherm populations"
bibliography: ./bib/refs.bib
csl: ./bib/ecology-letters.csl
format:
  docx:
    reference-doc: ./bib/template.docx
    output-file: "ms_revised.docx"
editor_options: 
  chunk_output_type: console
execute:
  freeze: auto  # re-render only when source changes
  cache: false
  echo: false
  warning: false
  error: false
  include: true
crossref:
  fig-title: 'Figure'
  fig-labels: arabic
  title-delim: "-"
  fig-prefix: "Figure"
  tbl-prefix: "Table"
editor: 
  markdown: 
    wrap: 72
---

@@

## Detailed trait category analyses for lnRR

  # Filter out antioxidant and NAs. For terrestrial, need to filter out endurance and max_MR because only 3 effect sizes
  lnRRQ10_data_m <- lnRRQ10_data %>% filter(habitat == "m" & !trait_category =="antiox") %>% filter(!is.na(c_accl_time))
  lnRRQ10_data_f <- lnRRQ10_data %>% filter(habitat == "f" & !is.na(c_accl_time))
  lnRRQ10_data_t <- lnRRQ10_data %>% filter(habitat == "t" & !is.na(c_accl_time)) %>% filter(!trait_category %in% c("endurance", "max_MR", "other"))

@@

## Detailed trait category analyses for lnCVR

@@

  } else{
#########################
  # Detailed Trait Categories
  # Models for the supplement
#########################
        model2.3 <- readRDS("./output/models/lnCVRQ10_2.3")
        model2.3f <- readRDS("./output/models/lnCVRQ10_model2.3f")
        model2.3m <- readRDS("./output/models/lnCVRQ10_model2.3m")
        model2.3t <- readRDS("./output/models/lnCVRQ10_model2.3t")
  model2.3_bayes <- readRDS("./output/models/lnCVRQ10_2.3_bayes")
}
```

>>>>

---
title: "Limited plasticity but increased variance in physiological rates across ectotherm populations under climate change"
bibliography: ./bib/refs.bib
csl: ./bib/ecology-letters.csl
format:
  docx:
    reference-doc: ./bib/template.docx
    output-file: "ms_revised.docx"
editor_options: 
  chunk_output_type: console
execute:
  freeze: auto  # re-render only when source changes
  cache: false
  echo: false
  warning: false
  error: false
  include: true
crossref:
  fig-title: 'Figure'
  fig-labels: arabic
  title-delim: "-"
  fig-prefix: "Figure"
  tbl-prefix: "Table"
editor: 
  markdown: 
    wrap: 72
---

@@

##########################################
## lnRR - Detailed trait category analyses 
##########################################
  # Filter out antioxidant and NAs. For terrestrial, need to filter out endurance and max_MR because only 3 effect sizes
  lnRRQ10_data_m <- lnRRQ10_data %>% filter(habitat == "m" & !trait_category =="antiox") %>% filter(!is.na(c_accl_time))
  lnRRQ10_data_f <- lnRRQ10_data %>% filter(habitat == "f" & !is.na(c_accl_time))
  lnRRQ10_data_t <- lnRRQ10_data %>% filter(habitat == "t" & !is.na(c_accl_time)) %>% filter(!trait_category %in% c("endurance", "max_MR", "other"))

@@

##########################################
# lnCVR - Detailed trait analysis
##########################################

@@

  } else{
#########################
  # Detailed Trait Categories
  # Models for the supplement
#########################
        model2.3 <- readRDS("./output/models/lnCVRQ10_2.3")
        model2.3f <- readRDS("./output/models/lnCVRQ10_model2.3f")
        model2.3m <- readRDS("./output/models/lnCVRQ10_model2.3m")
        model2.3t <- readRDS("./output/models/lnCVRQ10_model2.3t")
  model2.3_bayes <- readRDS("./output/models/lnCVRQ10_2.3_bayes")
  }

##########################################
## lnVR - Detailed trait category analyses
##########################################
  # Filter out antioxidant and NAs. For terrestrial, need to filter out endurance and max_MR because only 3 effect sizes
  lnVRQ10_data_m <- lnVRQ10_data %>% filter(habitat == "m" & !trait_category =="antiox") %>% filter(!is.na(c_accl_time))
  lnVRQ10_data_f <- lnVRQ10_data %>% filter(habitat == "f" & !is.na(c_accl_time))
  lnVRQ10_data_t <- lnVRQ10_data %>% filter(habitat == "t" & !is.na(c_accl_time)) %>% filter(!trait_category %in% c("endurance", "max_MR", "other"))

  # Prune tree
   tree_m_VR <- tree_checks(lnVRQ10_data_m, tree = tree_lnVR, dataCol = "species_full", type = "prune")
   tree_f_VR <- tree_checks(lnVRQ10_data_f, tree = tree_lnVR, dataCol = "species_full", type = "prune")
   tree_t_VR <- tree_checks(lnVRQ10_data_t, tree = tree_lnVR, dataCol = "species_full", type = "prune")

   # cor matrices
   m_phylo_VR <- vcv(tree_m_VR, corr = TRUE)
   f_phylo_VR <- vcv(tree_f_VR, corr = TRUE)
   t_phylo_VR <- vcv(tree_t_VR, corr = TRUE)

if(rerun){
#########################
  # Detailed Trait Categories
  # Models for the supplement
#########################
  # More realistic model 
  model3.3 <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_category + type:trait_category, 
                                   V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c| spp,
                                            ~1 | trait,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_VR), dfs = "contain", 
                              test = "t", data = lnVRQ10_data)

   model3.3_bayes <- brms::brm(effect_size | se(sqrt(V), sigma = TRUE) ~  max_temp_c + c_accl_time + type + trait_category + type:trait_category + 
                                 (~1 | record_num) + 
                                 (1 | gr(species_full, cov = A_VR)) + 
                                 (1 + max_temp_c| spp) + 
                                 (1|trait), 
                               data2 = list(A_VR=A_VR), family = gaussian(), 
                               prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), 
                               control = list(adapt_delta = 0.98, max_treedepth = 12), 
                               chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 4, data = lnVRQ10_data)

  # Save models
  saveRDS(model3.3, file = "./output/models/lnVRQ10_3.3")
  saveRDS(model3.3_bayes, file = "./output/models/lnVRQ10_3.3_bayes")

  # Models for plotting

  model3.3f <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_category + type:trait_category, 
                                  V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c| spp,
                                            ~1 | trait,
                                            ~1 + trait_category | obs),
                              R = list(species_full = f_phylo_vr), dfs = "contain", 
                              test = "t", struct = c("GEN", "HCS"), phi = 0, data = lnVRQ10_data_f)

  saveRDS(model3.3f, file = "./output/models/lnVRQ10_model3.3f")

  model3.3m <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_category + type:trait_category, 
                                   V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c| spp,
                                            ~1 | trait,
                                            ~1 + trait_category| obs),
                              R = list(species_full = m_phylo_vr), dfs = "contain", 
                              test = "t", struct = c("GEN", "HCS"),  phi = 0, data = lnVRQ10_data_m)
  saveRDS(model3.3m, file = "./output/models/lnVRQ10_model3.3m")

  # Note that we need to drop interaction for terrestrial as it's not able to estimate. 
  model3.3t <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_category, 
                                   V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c | spp,
                                            ~1 | trait,
                                            ~1 + trait_category | obs),
                              R = list(species_full = t_phylo_vr), dfs = "contain", 
                              test = "t", struct = c("GEN", "HCS"),  phi = 0, data = lnVRQ10_data_t)
  saveRDS(model3.3t, file = "./output/models/lnVRQ10_model3.3t")

  } else{
#########################
  # Detailed Trait Categories
  # Models for the supplement
#########################
        model3.3 <- readRDS("./output/models/lnCVRQ10_3.3")
        model3.3f <- readRDS("./output/models/lnCVRQ10_model3.3f")
        model3.3m <- readRDS("./output/models/lnCVRQ10_model3.3m")
        model3.3t <- readRDS("./output/models/lnCVRQ10_model3.3t")
  model3.3_bayes <- readRDS("./output/models/lnCVRQ10_3.3_bayes")
}
```
